Large-scale harmonic patterns clustering

### 2010-02-26 Paper/Paper.tex
improved

\section{Previous Work}\label{sec:prevwork}
The idea is closely related to the \textit{Shingles} described by
Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. 
LSH implementation: \cite{E2LSH}.
Barrington et al. recently studied music texture using a set of ordered
MFCC vectors \cite{Barrington2009a}. 
Nearest neighbor for music also includes \cite{Cano2004} and ...

@@

\subsection{Echo Nest features}
Echo Nest analyze API: \cite{EchoNest}.

@@

\subsection{uspop2002}
uspop: \cite{uspop2002}.

@@

Vector quantization. Online learning.

@@

\begin{algorithm}
\caption{Pseudocode of Vector Quantization}
\label{alg:vq}
\begin{algorithmic}
\STATE$l$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < l \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
\STATE$c \leftarrow c + (p - c) * l$
\ENDFOR
\ENDFOR
\RETURN $u$
\end{algorithmic}
\end{algorithm}

@@

>>>>

\section{Previous Work}\label{sec:prevwork}
The idea is closely related to the \textit{Shingles} described by
Slaney and Casey \cite{Casey2006,Casey2007,Casey2008}. 
LSH: \cite{Datar2004}, LSH implementation: \cite{E2LSH}.
Barrington et al. recently studied music texture using a set of ordered
MFCC vectors \cite{Barrington2009a}. 
Nearest neighbor for music also includes \cite{Cano2004} and ...

@@

\subsection{Echo Nest features}
We use the Echo Nest analyze API \cite{EchoNest} throughout this work.
The API gives us a chroma vector (length $12$) for every music event, 
or ``segment'', for any song we upload to their platform. 
We also get an estimate of where the beats and the bars are. 
We transform the chroma vectors per segment into per beat vector using a 
simple average. We can then stack beat vectors for a number of bars (usually 1 or 2). 
We get a fixed-size patch by resampling the number of chroma vectors. Typically,
there are $4$ beats per bar. Thus an appropriate patch size for one bar
is $12$x$4$ or $12$x$8$.

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate. Practise showed us that they are reasonable, and the
size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Section \ref{sec:experiments}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{code}
\end{center}
\caption{{A typical code from a code book of size $200$. Patterns represented
$2$ bars and the pattern length was set to $16$.}}
\label{fig:code}
\end{figure}

@@

\subsection{uspop2002}
We had acces to a low quality (32kbps) version of the songs in the uspop 2002 data
set \cite{uspop2002}.
This well-known data show a great diversity of pop songs, but is not particular
otherwise. We also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
not contain any bar.

@@

Vector quantization \cite{Gersho1991}. Online learning.

@@

\begin{algorithm}
\caption{Pseudocode of Vector Quantization}
\label{algo:vq}
\begin{algorithmic}
\STATE$l$ learning rate
\STATE$\{P_n\}$ set of patches
\STATE$\{C_k\}$ codebook of $K$ codes
%\STATE $m \leftarrow b$
\REQUIRE $0 < l \leq 1$
\FOR{$nIters$}
\FOR{$p \in \{P_n\}$}
\STATE$c \leftarrow min_{c \in C_k} dist(p,c)$
\STATE$c \leftarrow c + (p - c) * l$
\ENDFOR
\ENDFOR
\RETURN $\{C_k\}$
\end{algorithmic}
\end{algorithm}

@@

\subsection{Setting}
We take one or two bars, normalize the patches to size 4, 8, or 16.
We roll the patch to be invariant to the key (on the patch level, not on
the song level). We learn a codebook of size $K$ over the cowbell dataset 
using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
is $1e-2$ for $20$ iterations over the whole dataset.

We use the codebook to encode our testing set (see Section \ref{sec:data}).
Each pattern is encoded with only one code. We can measure the average
distance between a pattern and its encoding. We can also measure the use
of the different codes.

\subsection{Pattern length}

\subsection{Visualization}

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes_lle}
\end{center}
\caption{{LLE visualization of the code book.}}
\label{fig:lle}
\end{figure}

\subsection{Specific Examples}
